---
title: "MACAGBA, JAN EDWARD F.-FA1"
output: pdf_document
date: "2024-02-01"
---

```{r skewness, include=TRUE}
library(e1071)
library(tinytex)

results=read.csv("C:/Users/asus/Documents/Macagba_R/data.csv",header=T)
results

skewness(results$arch1, na.rm = T)
skewness(results$arch2, na.rm = T)
skewness(results$prog1, na.rm = T)
skewness(results$prog2, na.rm = T)
```
In the data it would be seen that all of the examination subjects are all 
negative except for arch2 which is positive. It can be also seen that the prog1
and prog2 are closer to each other while arch1 is the one skewed to the very
left.

------------------------------------------------------------------------------

```{r skewness with Pearsons Equation, include=TRUE}
skew=function(x){
  mean=mean(x, na.rm = T)
  median=median(x, na.rm = T)
  sd=sd(x, na.rm = T)
  
  skewness=3*(mean-median)/sd
  
  return(skewness)
}

arch1_skewness=skew(results$arch1)
arch2_skewness=skew(results$arch2)
prog1_skewness=skew(results$prog1)
prog2_skewness=skew(results$prog2)

print(arch1_skewness)
print(arch2_skewness)
print(prog1_skewness)
print(prog2_skewness)
```
When using Pearsons equation, the data increased especially prog1 from a -0.329
it became a -0.643 the other data increased by 1. I would say that using
Pearsons equation has risk to it because there would be inaccuracies in the 
data.

------------------------------------------------------------------------------

```{r Stem and Leaf, include=TRUE}
Females=c(57,59,78,79,60,65,68,71,75,48,51,55,56,41,43,44,75,78,80,81,83,83,85)
Males=c(48,49,49,30,30,31,32,35,37,41,86,42,51,53,56,42,44,50,51,65,67,51,56,
        58,64,64,75)

stem(Females)
stem(Males)
```
The advantages of using the stem and leaf display is that for small data it 
would be much easier to find out the precise data instead of using a histogram.
A stem and leaf would be easier to create than a histogram and also you would 
be seeing the individual data and not just different bars.

------------------------------------------------------------------------------

```{r boxplot, include=TRUE}
Females=c(57,59,78,79,60,65,68,71,75,48,51,55,56,41,43,44,75,78,80,81,83,83,85)
Males=c(48,49,49,30,30,31,32,35,37,41,86,42,51,53,56,42,44,50,51,65,67,51,56,
        58,64,64,75)

boxplot(Females, Males, names=c("Females", "Males"), main="Grades by Gender", 
        ylab="Grades",col=c("#FFE5E5","#756AB6"))
```
The finding for the boxplot is that the grades for the Female group is higher 
than the Male group with the median of the Female group being higher.  
